Active Contextual Entropy Search

### 2016-05-31 2016_ICML/poster/content/left_column.tex
Poster content

\begin{block}{\blocktitle{Background}}
\textbf{Contextual policy search} \cite{deisenroth_survey_2013} aims at learning an upper-level policy $\pi_\omega(\theta \vert s)$, where
\begin{itemize}
 \item $s \in \mathbb{R}^{n_s}$ encodes the current context (task)
 \item $\theta \in \mathbb{R}^{n_\theta}$ is a parameter vector for a lower-level policy such as a DMP \cite{ijspeert_dynamical_2013}
 \item $\omega \in \mathbb{R}^{n_\omega}$ is the parameter vector of the upper-level policy to be learned
\end{itemize}
such that the expected return over all contexts $J_\omega = \int_s \mu(s) \int_\theta \pi_\omega(\theta \vert s) R(s, \theta) ds d\theta$ is maximized.

One class of contextual policy search methods are \textbf{gradient-free, local-search methods} such as cost-regularized kernel regression \cite{Kober2012} and contextual relative entropy search (C-REPS) \cite{kupcsik_data-efficient_2013}.

\textbf{Bayesian optimization for contextual policy search (BO-CPS)} \cite{metzen_bayesian_2015} is an alternative model-based global-search approach, which is very efficient for low-dimensional search spaces combined with an expensive cost function.
\begin{itemize}
 \item Use surrogate model (Gaussian process) to model expected return for context-parameter pairs (allows generalization over similar contexts)
 \item Acquisition function: GP upper confidence bound (GP-UCB) \cite{srinivas_gaussian_2010}, entropy search \cite{hennig_entropy_2012}
 \item Perform optimization (DIRECT + L-BFGS) of acquisition function over parameter space for fixed context (context determined by the environment or chosen randomly)
\end{itemize}

@@

\begin{block}{\blocktitle{Active Contextual Entropy Search}}
\emph{Idea:} Extend entropy search such that it allows selecting \emph{actively} the context for the next trial, in which it expects to learn the most

@@

Active Contextual Entropy Search (ACES):
\begin{itemize}
 \item Let $p_{max}(\theta \vert s)$ denote the CPD that maximum return given context $s$ is at $\theta$
 \item Loss: expected change of relative entropy in context $s$ after performing a trial in context $s^q$with parameter $\theta^q$: $\mathcal{L}^s(s^q, \theta^q) = \mathcal{L}(p_{max}[s^q,\theta^q](\theta \vert s)) - \mathcal{L}(p_{max}(\theta \vert s))$
 \item $s^q, \theta^q = \arg\min_{(s^q, \theta^q)} \mathcal{L}^{s^q}(s^q,
\theta^q)$ would not account for information gained about the
optima in contexts $s \neq s^q$
 \item Acquisition function: $\text{ACES}(s^q, \theta^q) = \sum_{i=1}^{N_s}
\mathcal{L}^{s^c_i}(s^q, \theta^q)$, trial contexts $s^c_i \sim \mathcal{U}(S)$
 \item Evaluating $\text{ACES}(s^q, \theta^q)$ is expensive but GP has intrinsic length scales; \\approximate $\text{ACES}(s^q, \theta^q) \approx \sum_{s \in \text{NN}(s^q,
\{s^c\}, N_{nn})}  \mathcal{L}^s(s^q, \theta^q)$
 \item NN returns the $N_{nn}$ nearest neighbors of  $s^q$ in $\{s^c\}$ according to a Mahalanobis distance with the (anisotropic) GP length scales on the diagonal of the covariance
 \item Small $N_{nn}$ is faster but approximates $\text{ACES}(s^q, \theta^q)$ using only local information (local in context space), we compare $N_{nn}=1$ and $N_{nn}=20$
 \item Optimization based on CMA-ES (noisy acquisition function)
\end{itemize}
\end{block}

@@

\begin{figure}
\centering
\includegraphics[width=0.4\columnwidth]{../pics/regret_illustration}
\includegraphics[width=0.4\columnwidth]{../pics/acq_comparison}
\caption{(Left) Illustration of GP posterior, probability of maximum $p^\star$, expected regret ER, and  $\mathbb{E}[\max(f(\mathbf{x}) - f(1.5), 0)]$ (scale on the right-hand side). (Right) Illustration of GP posterior and different acquisition function. Absolute values have been normalized such that the mean value of an acquisition function is $0.5$. Best seen in color.}
\label{fig:MRS_illustration}
\end{figure}

>>>>

\begin{block}{\blocktitle{Background}}
\textbf{Bayesian optimization} \cite{shahriari_taking_2016}:
\begin{itemize}
\item aims at black-box optimization problems $\mathbf{x} = \arg\min_{\mathbf{x} \in \mathcal{X}} f(x)$ of some function $f: \mathcal{X} \to \mathbb{R}$ on some bounded set $\mathcal{X} \subset \mathbb{R}^D$.
\item maintains a \emph{probabilistic model} for $f(\mathbf{x})$, typically a Gaussian process (GP)
\item decides on a query point $\mathbf{x}_{n+1}$ where $f$ will be evaluated next based on GP posterior of first $n$ observations $\mathcal{D}_n=\{(\mathbf{x}_i, y_i)\}_{i=1}^n$ and an \emph{acquisition function}
\end{itemize}

Acquisition functions decide where to query next:
\begin{itemize}
\item Upper-Confidence Bound (UCB): $a_{UCB}(\mathbf{x};
\mathcal{D}_n) = \mu_{n}(\mathbf{x}) + \kappa_n \sigma_{n}(\mathbf{x})$
\end{itemize}

@@

\begin{block}{\blocktitle{Minimum Regret Search}}
\emph{Idea:}

@@

Minimum Regret Search (MRS):
\begin{itemize}
 \item 
\end{itemize}
\end{block}

@@

\begin{figure}
\centering
\includegraphics[width=0.48\columnwidth]{../pics/regret_illustration}
\includegraphics[width=0.48\columnwidth]{../pics/acq_comparison}
\caption{(Left) Illustration of GP posterior, probability of maximum $p^\star$, expected regret ER, and  $\mathbb{E}[\max(f(\mathbf{x}) - f(1.5), 0)]$ (scale on the right-hand side). (Right) Illustration of GP posterior and different acquisition function. Absolute values have been normalized such that the mean value of an acquisition function is $0.5$. Best seen in color.}
\label{fig:MRS_illustration}
\end{figure}
\end{block}

\begin{block}{\blocktitle{Synthetic Single-Task Benchmark}}
\begin{itemize}
 \item Comparison of different approaches for exploration
 \begin{enumerate}
  \item Active contextual entropy search ($N_{nn} = 20$ and $N_{nn} = 1$)
  \item BO-CPS + Entropy Search +  random context selection
  \item BO-CPS + GP-UCB with $\kappa=5.0$ (``pure exploration'') + random context selection
  \item Random parameter and context selection
 \end{enumerate}
 \item Entropy Search outperforms GP-UCB because UCB aggressively samples parameters at the boundaries (high uncertainty but low global information)
 \item ACES outperforms random context selection, in particular when $N_{nn} = 20$
 \item ACES\_20 tends to avoid performing trials at the boundaries of the context space
\end{itemize}

### 2016-05-31 2016_ICML/poster/content/right_column.tex
Poster content

%place your content here, feel free to create subcontent files
\begin{block}{\blocktitle{Experimental Setup}}
\begin{center}
\raisebox{0.75cm}{
\includegraphics[width=.2\columnwidth]{pics/throw}}
\hspace*{1cm}
%\includegraphics[width=.35\columnwidth]{pics/xy_scatter_color_tau}
\hspace*{1cm}
%\includegraphics[width=.35\columnwidth]{pics/xy_scatter_color_g0}
\begin{itemize}
 \item Ball-throwing in simulation of COMPI arm (\url{github.com/rock-simulation/mars})
 \item Contexts chosen from target area: $[1, 2.5]m \times [-1, 1]m$
 \item Manually initialized joint-space DMP such that a fixed target position is hit
 \item Adapt to new target by modifying meta-parameters (execution time $\tau$ and angle of first joint $g_0$); $\tau \in [0.2, 5]$ and $g_0 \in \left[-\frac{\pi}{2}, \frac{\pi}{2}\right]$
 \item $R(s, \theta) = -\vert\vert s - b_s(\theta) \vert\vert_2^2 - 0.01\sum_t v_t^2$
 \begin{itemize}
   \item $b_s(\theta)$:  position hit for parameters $\theta$; $v_t$: joint velocities at time $t$
 \end{itemize}
 \item Deterministic affine upper-level policy; trained with C-REPS \cite{kupcsik_data-efficient_2013} update on all samples
\end{itemize}
\end{center}
\end{block}

\begin{block}{\blocktitle{Results}}
\begin{center}
\begin{itemize}
 \item Comparison of different approaches for exploration
 \begin{enumerate}
  \item Active contextual entropy search ($N_{nn} = 20$ and $N_{nn} = 1$)
  \item BO-CPS + Entropy Search +  random context selection
  \item BO-CPS + GP-UCB with $\kappa=5.0$ (``pure exploration'') + random context selection
  \item Random parameter and context selection
 \end{enumerate}
 \item Entropy Search outperforms GP-UCB because UCB aggressively samples parameters at the boundaries (high uncertainty but low global information)
 \item ACES outperforms random context selection, in particular when $N_{nn} = 20$
 \item ACES\_20 tends to avoid performing trials at the boundaries of the context space
\end{itemize}

\begin{figure}
\centering
\includegraphics[width=.9\columnwidth]{../pics/empirical_comparison} \\
\includegraphics[width=.9\columnwidth]{../pics/hist}
\caption{(Top) Median and mean simple regret over $250$ repetitions for different acquisition functions. Shown is the simple regret of the recommendation $\mathbf{\tilde x}_N$ after $N$ trials, i.e., the point which maximizes the GP posterior
mean. (Bottom) Histogram of the simple regret after performing $N=100$ trials for different acquisition functions (note the log-scales).}
\label{fig:empirical_comparison}
\end{figure}

@@

@@

\begin{block}{\blocktitle{Outlook}}
\begin{itemize}
 \item Comparison to other active task (context) selection approaches \cite{fabisch_accounting_2015}
 \item Scaling up to higher dimensions using approaches such as REMBO
\cite{wang_bayesian_2013}
 \item Evaluation on real systems and related tasks such as robotic locomotion \cite{calandra_bayesian_2014,lizotte_automatic_2007} and robot grasping \cite{kroemer_combining_2010}
 \item More information on project BesMan (Behaviors for Mobile Manipulation):
\url{http://robotik.dfki-bremen.de/en/research/projects/besman.html}
\end{itemize}
\end{block}

>>>>

%place your content here, feel free to create subcontent files

\begin{block}{\blocktitle{Results}}
\begin{center}

\begin{figure}
\centering
\includegraphics[width=.8\columnwidth]{../pics/empirical_comparison} \\
\includegraphics[width=.8\columnwidth]{../pics/hist}
\caption{\emph{No model mismatch:} Simple regret of the recommendation $\mathbf{\tilde x}_N$ after $N$ trials over $250$ repetitions for different acquisition functions.}
\label{fig:empirical_comparison}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=.8\textwidth]{../pics/empirical_comparison_mm} \\
\includegraphics[width=.8\textwidth]{../pics/hist_mm}
\caption{\emph{Model mismatch:} Simple regret of the recommendation $\mathbf{\tilde x}_N$ after $N$ trials over $250$ repetitions for different acquisition functions.}
\label{fig:empirical_comparison_mm}
\end{figure}

@@

\begin{block}{\blocktitle{Outlook}}
\begin{itemize}
 \item For multi-task learning and results on simulated robotic task: see paper
 \item Open questions: treatment of GP hyperparameters, more efficient approximation techniques for MRS, regret bounds for MRS?
\end{itemize}
\end{block}
